---
title: "Exploratory Data Analysis - Job Offers Dataset (Remotive + Jobicy)"
author: "Ramiro Crespi"
format: html
editor: visual
---

Load Libraries

```{r}
library(tidyverse)
library(modelr)
library(naniar)
library(tidytext)
library(wordcloud)
library(tm)
    
```

Load data

```{r}
setwd('/Users/user/Documents/GitHub/job-data-project')
df <- read_csv('data/processed/all_jobs.csv')

```

### Initial Analysis

```{r}
glimpse(df)
summary(df)
sapply(df, n_distinct)

df %>% count(region)
df %>% count(seniority)
df %>% count(category, sort = TRUE)


```

Filtering 'Other' Values and Extreme Salaries

```{r}
other_seniorities <- df %>% filter(seniority == "Other")
other_categories <- df %>% filter(category == "Other")
other_regions <- df %>% filter(region == "Other")

df_clean <- df %>%
  filter(salary_cleaned < 1e6)

quantile(df$salary_cleaned, probs = seq(0.9, 1, 0.01), na.rm = TRUE)


```

### Salary Visualizations

```{r}

ggplot(df, aes(x = salary_cleaned)) +
  geom_histogram(bins = 30, fill = "#1f77b4") +
  labs(title = "Distribución de salarios (USD)", x = "Salario", y = "Frecuencia") +
  xlim(30000, 300000) +
  theme_minimal()

ggplot(df, aes(x = salary_cleaned)) +
  geom_histogram(aes(y = ..density..), fill = "#1f77b4") +
  geom_density(color = "firebrick", linewidth = 1, bw = 9000) +
  labs(title = "Distribución de salarios (USD)", x = "Salario", y = "Densidad") +
  xlim(3000, 300000) +
  theme_minimal()

ggplot(df_clean, aes(x = category, y = salary_cleaned)) +
  geom_boxplot(fill = "orange") +
  coord_flip() +
  labs(title = "Salary Distribution by Category", x = "Category", y = "Salary") +
  theme_minimal()

ggplot(df_clean, aes(x = region, y = salary_cleaned)) +
  geom_boxplot(fill = "forestgreen") +
  coord_flip() +
  labs(title = "Salary Distribution by Region", x = "Region", y = "Salary") +
  theme_minimal()

ggplot(df_clean, aes(x = seniority, y = salary_cleaned)) +
  geom_boxplot(fill = "steelblue") +
  coord_flip() +
  labs(title = "Salary Distribution by Seniority", x = "Seniority", y = "Salary") +
  theme_minimal()

```

### Categories and Seniorities Analysis

```{r}
df %>%
  count(category, sort = TRUE) %>%
  ggplot(aes(x = reorder(category, n), y = n)) +
  geom_col(fill = "purple") +
  coord_flip() +
  labs(title = "Cantidad de Ofertas por Categoría", x = "Categoría", y = "Cantidad") +
  theme_minimal()

df %>%
  count(seniority, sort = TRUE) %>%
  ggplot(aes(x = reorder(seniority, n), y = n)) +
  geom_col(fill = "dodgerblue") +
  coord_flip() +
  labs(title = "Cantidad de Ofertas por Seniority", x = "Seniority", y = "Cantidad") +
  theme_minimal()

```

### Time evolution

```{r}
df %>%
  mutate(fecha = as_date(pubDate)) %>%
  count(fecha) %>%
  ggplot(aes(x = fecha, y = n)) +
  geom_line(color = "firebrick", size = 1.1) +
  labs(title = "Job Postings Over Time", x = "Date", y = "Count") +
  theme_minimal()

df %>%
  mutate(pubDate = as.Date(pubDate),
         day = weekdays(pubDate)) %>%
  count(day) %>%
  ggplot(aes(x = reorder(day, n), y = n)) +
  geom_col(fill = "darkred") +
  labs(title = "Job Postings by Weekday", x = "Day", y = "Count") +
  coord_flip() +
  theme_minimal()

```

Subsets by Category (Data, Engineering, Other)

```{r}
categories <- list("Data", "Engineering", "Other")

for (cat in categories) {
  df_cat <- df %>% filter(category == cat)

  print(
    ggplot(df_cat, aes(x = salary_cleaned)) +
      geom_histogram(bins = 30, fill = "#1f77b4") +
      labs(title = paste("Salary Distribution -", cat),
           x = "Salary", y = "Frequency") +
      xlim(30000, 300000) +
      theme_minimal()
  )
}

```

Wordclouds by Category

```{r}
for (cat in categories) {
  df %>%
    filter(category == cat) %>%
    pull(jobTitle) %>%
    tolower() %>%
    wordcloud::wordcloud(max.words = 100)
}

```

Category vs Region

```{r}
df %>%
  count(region, category) %>%
  ggplot(aes(x = reorder(region, n), y = n, fill = category)) +
  geom_col(position = "fill") +
  coord_flip() +
  labs(title = "Job Offers by Region and Category") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1))

```

### Statistical (ANOVA) Analysis

```{r}
# ANOVA for salary by category
anova_cat <- aov(salary_cleaned ~ category, data = df_clean)
summary(anova_cat)

# ANOVA for salary by region
anova_reg <- aov(salary_cleaned ~ region, data = df_clean)
summary(anova_reg)

# ANOVA for salary by seniority
anova_sen <- aov(salary_cleaned ~ seniority, data = df_clean)
summary(anova_sen)

```

### Final Notes

```{r}
cat("Seniority level does not show significant differences in salary by itself. However, it is likely to have a meaningful effect when segmented by region, category, or both. This suggests that there may be interaction effects worth exploring in further analyses.")

```

Interacciones

```{r}
# Interaction Models

# Seniority * Region
interaction_model_1 <- lm(salary_cleaned ~ seniority * factor(region), data = df_clean)
summary(interaction_model_1)

# Seniority * Category
interaction_model_2 <- lm(salary_cleaned ~ seniority * factor(category), data = df_clean)
summary(interaction_model_2)

# Region * Category
interaction_model_3 <- lm(salary_cleaned ~ region * category, data = df_clean)
summary(interaction_model_3)

# Full 3-way interaction
interaction_model_full <- lm(salary_cleaned ~ seniority * region * category, data = df_clean)
summary(interaction_model_full)

```

```{r}
anova(interaction_model_full)
```

### 🔎 Summary of Interaction Effects on Salary

A series of linear models were fitted to explore whether the effect of `seniority` on salary depends on `region`, `category`, or both. The models tested interaction effects between pairs and across all three variables.

Model 1: seniority × region

```{r}
anova(interaction_model_1)

```

-   **Region** is the most influential predictor (p \< 2e-16).

-   **Seniority** has a small but statistically significant effect (p = 0.036).

-   The interaction between **seniority and region** is **marginally significant** (p = 0.091), suggesting potential combined effects on salary worth exploring further.

    Model 2: seniority × category

    ```{r}
    anova(interaction_model_2)
    ```

-   **Category** is weakly significant (p = 0.042).

-   **Seniority** alone is **not** significant (p = 0.169).

-   The interaction between **seniority and category** is **not significant** (p = 0.614), indicating no clear interaction effects in this model.

    Model 3: region × category

    ```{r}
    anova(interaction_model_3)
    ```

-   **Region** again emerges as a highly significant predictor (p \< 2e-16).

-   **Category** is not significant (p = 0.153).

-   The interaction **region × category** is significant (p = 0.023), suggesting salaries vary across region–category combinations.

Full Model: seniority × region × category

```{r}
anova(interaction_model_full)
```

-   Only **region** and **seniority** are significant main effects (p \< 2e-16 and p = 0.035, respectively).

-   All interaction terms, including the **three-way interaction** (`seniority:region:category`), are **not statistically significant**.

-   The marginal significance of the **seniority × region** interaction (p = 0.090) is preserved here, reinforcing the idea that this interaction may deserve further attention.

### 💡 Interpretation

These results confirm that:

-   **Region consistently plays a key role in salary differences**.

-   There is **some evidence that seniority interacts with region**, but not with category.

-   The **full interaction model adds complexity without significantly improving explanation**, suggesting a simpler model (e.g., `seniority * region`) may suffice.

```{r}
library(plotly)

p <- df_clean %>%
  filter(!is.na(salary_cleaned), category != "Other", region != "Other") %>%
  ggplot(aes(x = category, y = salary_cleaned, fill = region, text = paste("Region:", region, "<br>Salary:", salary_cleaned))) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) +
  coord_flip() +
  labs(title = "Salaries per Category and Region", x = "Category", y = "Salary (USD)") +
  theme_minimal()

ggplotly(p, tooltip = "text")

```

```{r}
p2 <- df_clean %>% filter(!is.na(salary_cleaned), region != "Other", seniority != "Other") %>%
  ggplot(aes(x = seniority, y = salary_cleaned, fill = region, text = paste("Region:", region, "<br>Seniority:", seniority, "<br>Salary:", salary_cleaned))) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.8)) +
  labs(title = "Salaries by Region and Seniority Level", x = "Seniority", y = "Salary (USD)") +
  theme_minimal()

ggplotly(p2, tooltip = "text")

```
